Discrete Probability for Discrete Quantum Computing

### 2016-04-30 prob.tex
example dempster shafer

The probability measure above can be understood in reference to
Ex.~\ref{ex1} by mapping each numeric value to the smallest interval
containing it. The probability measure appears quite intuitive from
that perspective. However, if we attempt to understand the probability
measure independently of Ex.~\ref{ex1} we would hard pressed to
justify the following assignments:
\[ 
\pmeas(\{HH\}) = \unlikely, \quad\pmeas(\{TH\}) = \likely, \quad
\pmeas(\{HH\}\cup\{TH\}) = \necess 
\]
which assert that an \unlikely-event whose probability is in the range
$[0,\frac{1}{2}]$ and a \likely-event whose probability is in the
range $[\frac{1}{2},1]$ combine to a \necess-event whose probability
is in the range $[1,1]$! To understand this mystery, it is important
to investigate the mathematical definition of interval-valued
probabilities in detail.

In interval-valued probability measures, if the probability of an
event $E$ is $[a,b]$, we think of the left-endpoint~$a$ as
representing the strength of the evidence that supports $E$, and the
right-endpoint~$b$ as the strength of the evidence that contradicts
$E$. If we define $1-\left[a,b\right]=\left[1-b,1-a\right]$, then we
intuitively have
\begin{eqnarray*}
\pmeas\left(\emptyset\right) & = & \imposs\\
\pmeas\left(\Omega\right) & = & \necess\\
\pmeas\left(\Omega\backslash E\right) & = & 1-\pmeas\left(E\right)
\end{eqnarray*}
Next, if we define $\sum_{i}\left[a_{i},b_{i}\right]=\left[\sum_{i}a_{i},\sum_{i}b_{i}\right]$,
then 
\begin{itemize}
\item for a collection $E_{i}$ of pairwise disjoint events, we have $\pmeas\left(\bigcup_{i}E_{i}\right)\subseteq\sum_{i}\pmeas\left(E_{i}\right)$. 
\end{itemize}
Notice that the equality may not hold in general. This statement says
that the evidences of $\bigcup_{i}E_{i}$ is at least as strong as
putting all the evidences of $E_{i}$ together, but some evidence
may only be acquired for $\bigcup_{i}E_{i}$ as the whole. Therefore,
$\pmeas\left(\bigcup_{i}E_{i}\right)$ is a subset of $\sum_{i}\pmeas\left(E_{i}\right)$,
but may not equal. In our example, 
\begin{eqnarray*}
 &  & \pmeas(\{HH,TH\})=\necess=[1,1]\\
 & \subseteq & \left[\frac{1}{2},\frac{3}{2}\right]=\left[0,\frac{1}{2}\right]+\left[\frac{1}{2},1\right]=\unlikely+\likely=\pmeas(\{HH\})+\pmeas(\{TH\})\textrm{ .}
\end{eqnarray*}
However, $\pmeas(\{HH,TH\})$ is a proper subset of $\pmeas(\{HH\})+\pmeas(\{TH\})$
because if we check the complement of $\{HH,TH\}$, we have 
\begin{eqnarray*}
 &  & \pmeas(\{HH,TH\})=1-\pmeas(\{HT,TT\})=1-\imposs=\necess\textrm{ ,}
\end{eqnarray*}
and $\pmeas(\{HT,TT\})=\imposs$ cannot be used to reasoning the probability
of $\{HH\}$ and $\{TH\}$ individually.

>>>>

The probability measure above can be understood in reference to
Ex.~\ref{ex1} by mapping each numeric value to the smallest interval
containing it. The probability measure appears quite intuitive from
that perspective. However, we also must be able to understand the
probability measure independently of Ex.~\ref{ex1} and in particular
justify how combinations such as
\[ 
\pmeas(\{HH\}) = \unlikely, \quad\pmeas(\{TH\}) = \likely, \quad
\pmeas(\{HH\}\cup\{TH\}) = \necess 
\]
which assert that an \unlikely-event whose probability is in the range
$[0,\frac{1}{2}]$ and a \likely-event whose probability is in the
range $[\frac{1}{2},1]$ combine to a \necess-event whose probability
is in the range $[1,1]$! To understand this mystery, we look at
another more general example and then give the formal mathematical
definition of interval-valued probabilities.

\amr{See A Simple View of the Dempster-Shafer Theory of Evidence and 
  its Implication for the Rule of Combination Lotfi A. Zadeh, AAAI,
  1986}

\begin{example}[Dempster-Shafer Theory of Evidence] We have five
  employees whose precise age is not known. All is give is a range for
  each employee:
\[\begin{array}{rcl}
\textrm{Age}(M_1) &=& [22,26] \\
\textrm{Age}(M_2) &=& [20,22] \\
\textrm{Age}(M_3) &=& [30,35] \\
\textrm{Age}(M_4) &=& [20,22] \\
\textrm{Age}(M_5) &=& [28,30] 
\end{array}\]
What is the probability that an employee's age is in the range
$[20,25]$. Looking at the data, it is \emph{possible} for $M_1$'s age
to be within the range, it is \emph{not possible} for $M_3$'s age to
be within the range, and it is \emph{certain} or \emph{necessary} that
$M_4$'s age is in the range. Clearly saying that an event is
\emph{necessary} is equivalent to saying that its complement is
\emph{not possible}. Aggregating the results for the five employees,
we can calculate that is necessary that 2 employees have their ages
within the range, and it is possible that 3 employees have their ages
within the range. We express this formally as saying that the
probability of the event is $[0.4,0.6]$. Now consider another event
asking whether the ages are some other disjoint range
$[26,40]$. Reasoning in a similar way we calculate that the
probability for this event is $[0.4,0.6]$. Now let's take the two events
together and ask about the possibility of the ages to be in the range
$[20,40]$. Clearly that probability must be $[1,1]$ as every
employee's age is in that range. The way to calculate this probability
from the individual probabilities of the events is as follows: 
\begin{itemize}
\item \amr{can you try to fix this... perhaps it is not consistent
    with the interval-valued formalization but it is really nice if we
    can make it work}
\end{itemize}
\qed\end{example}

In interval-valued probability measures, if the probability of an
event $E$ is $[a,b]$, we think of the left-endpoint~$a$ as
representing the strength of the evidence that supports $E$, and the
right-endpoint~$b$ as the strength of the evidence that contradicts
$E$.

Thus if we have an event $E$ with probability $[a,b]$ where
$a=0.1$ and $b=0.7$, we have that:
\begin{itemize}
\item the strength of evidence supporting $E$ is 0.1; since either $E$
  or its complement must happen, we conclude that there is 0.9
  evidence supporting the complement of $E$; 
\item the strength of evidence contradicting $E$ is 0.7; again since
  either $E$ or its complement must happen, we conclude that there is
  0.3 evidence contradicting the complement of $E$.
\end{itemize}

Turning things around, the strength of evidence that contradicts
$E$ is evidence supporting the complement of $E$

The complement of $E$ must therefore have probability
$\left[1-b,1-a\right]$ which we abbreviate $1-\left[a,b\right]$:

\begin{eqnarray*}
\pmeas\left(\emptyset\right) & = & \imposs\\
\pmeas\left(\Omega\right) & = & \necess\\
\pmeas\left(\Omega\backslash E\right) & = & 1-\pmeas\left(E\right)
\end{eqnarray*}
Next, if we define $\sum_{i}\left[a_{i},b_{i}\right]=\left[\sum_{i}a_{i},\sum_{i}b_{i}\right]$,
then 
\begin{itemize}
\item for a collection $E_{i}$ of pairwise disjoint events, we have $\pmeas\left(\bigcup_{i}E_{i}\right)\subseteq\sum_{i}\pmeas\left(E_{i}\right)$. 
\end{itemize}
Notice that the equality may not hold in general. This statement says
that the evidences of $\bigcup_{i}E_{i}$ is at least as strong as
putting all the evidences of $E_{i}$ together, but some evidence
may only be acquired for $\bigcup_{i}E_{i}$ as the whole. Therefore,
$\pmeas\left(\bigcup_{i}E_{i}\right)$ is a subset of $\sum_{i}\pmeas\left(E_{i}\right)$,
but may not equal. In our example, 
\begin{eqnarray*}
 &  & \pmeas(\{HH,TH\})=\necess=[1,1]\\
 & \subseteq & \left[\frac{1}{2},\frac{3}{2}\right]=\left[0,\frac{1}{2}\right]+\left[\frac{1}{2},1\right]=\unlikely+\likely=\pmeas(\{HH\})+\pmeas(\{TH\})\textrm{ .}
\end{eqnarray*}
However, $\pmeas(\{HH,TH\})$ is a proper subset of $\pmeas(\{HH\})+\pmeas(\{TH\})$
because if we check the complement of $\{HH,TH\}$, we have 
\begin{eqnarray*}
 &  & \pmeas(\{HH,TH\})=1-\pmeas(\{HT,TT\})=1-\imposs=\necess\textrm{ ,}
\end{eqnarray*}
and $\pmeas(\{HT,TT\})=\imposs$ cannot be used to reasoning the probability
of $\{HH\}$ and $\{TH\}$ individually.
